\ \\ \LARGE\bf Rolling Horizon methods for Games with Continuous States and Actions

### 2014-04-11 lpaper/lunarcont.tex
added Michelle's fixes

\begin{abstract}
It is often the case that games have continuous dynamics and allow for continuous actions, possibly with with some added noise. For larger games with complicated dynamics, having agents learn offline behaviours in such a setting is a daunting task. On the other hand, provided a forward model is available, one might try to spread the cost of search/learning in a rolling horizon fashion (e.g. as in Monte Carlo Tree Search). In this paper we compare T-HOLOP (Truncated Hierarchical Open Loop Planning), an open loop planning algorithm at least partially inspired by MCTS with a version evolutionary planning using CMA-ES (which we call EVO-P) in two planning benchmark problems (Cart Balancing and the Double Integrator) and Lunar Lander, a classic arcade game. He show that EVO-P outperforms T-HOLOP in the classic benchmarks, while T-HOLOP is unable to find a solution using the same heuristic. We conclude that off-the-shelf evolutionary algorithms can be used successfully in a rolling horizon setting, and that a different type of heuristics might be needed under different optimisation algorithms.

@@

\section{Introduction} \label{Introduction}
One of the most common uses of function optimisation is for solving control problems. If an agent has access to a generative model of the world (i.e. a simulator) in which it is going to act, a control problem becomes a search/optimisation problem, usually referred to as ``simulation based planning''. Recently, a family of algorithms, mostly known as Monte Carlo Tree Search (MCTS)~\cite{browne2012survey}, has been applied to planning problems of discrete states and actions with considerable success. The focus of these algorithms is to attack extremely large state spaces of perfect information games, where one can sample rewards from the state space easily (e.g. Computer Go~\cite{gelly2006modification}). With access to a generative model (from which one can easily sample) and perfect sensor information, the following procedure works well: the agent receives sensor information, formulates a plan of action, performs the first action of that plan, receives a new state, re-formulates a plan, acts again (using the first action of its plan),  ad infinitum.

@@

Recently, algorithms stemming from, or at least partially inspired by, MCTS aiming at solving planning problems involving continuous states and actions have come to light. A typical such algorithm is Hierarchical Optimistic Open Loop Planing (HOLOP)~\cite{weinstein2012bandit}. HOLOP is based partially on a strong~\cite{hoo2011} real-valued optimisation algorithm called HOO (for ``Hierarchical Optimistic Optimisation''). In this paper two contributions are made; the first contribution is the comparison of two versions of a very popular and strong evolutionary strategy: EVO-P, the open loop planning algorithm using Covariance Matrix Adaptation Evolutionary Strategy (CMA-ES)~\cite{hansen2003reducing} with a version of HOLOP, which stands for Hierarchical Open Loop Optimistic Planning~\cite{hoo2011} on two standard benchmark problems, the double integrator and the reverse pendulum.  For speed purposed, we use T-HOO (Truncated Hierarchical Optimistic Optimisation), a faster version of HOO~\cite{hoo2011}, as the engine of HOLOP, hence we name the algorithm used T-HOLOP.  Our second contribution is the use of the these algorithms in Lunar Lander, a popular arcade game, coupled with a strong two-stage heuristic. We show that while EVO-P performs reasonably well, T-HOLOP struggles.

While evolution is commonly used to evolve a reactive neural network controller (in a process commonly called neuro-evolution) for these and other RL problems,
it should be noted that evolution is being applied in a very different way here. In this paper, evolution is applied to perform each action given the current state.  This approach can only be applied when a generative
model is available, but has the advantage of offering immediate good performance without any
prior learning.  The disadvantage, compared to neuro-evolution, is that every action performed requires
CPU time for the simulations.  This is described in more detail in section \ref{sec:Planning} and in line with what is being done in Perez et al. \cite{perez2013rolling}, although this time in a continuous setting.

@@

This is the first background section of this paper, which aims to introduce the two core optimisation algorithms on top of which our contributions are made. We will present both algorithms from an implementer's viewpoint, without explicitly justifying the design and mathematical choices behind them, nor going into too much detail. Interested readers should look at the cited bibliography for extensive discussions.

In general, the optimisation problem can be defined as: Given a function $f:\bm{X}\rightarrow Y $, where $X \in \mathcal{R}^n$, with $n$ being the length of vector $\bm{X}$. The domain $\bm{X}$ is usually called the search space. Our goal is to find a vector $\bm{x}_0$, for which $f(\bm{x}) > f(\bm{x_0}) $ for all $x$ (also known as ``minimisation'') or $f(\bm{x}) < f(\bm{x_0})$ for all $x$ (known as ``maximisation'').
There are an enormous number of algorithms devoted to this problem, but only two will be presented and compared here, which are both considered state of the art in their respective fields and do not require the calculation of a gradient. If one has an analytic expression of the gradient or if the functions proposed are smooth enough, it might be better to look elsewhere for solutions (e.g. Conjugate Gradient Methods~\cite{shewchuk1994introduction}).

@@

CMA-ES creates a population of individuals by sampling from an MND: $\mathcal{N}(m,C)$, which is uniquely defined by the distribution mean $m \in \mathcal{R}^{n}$ and its covariance matrix $C \in \mathcal{R}^{n \times n}$. The top of the density function, which corresponds to $m$, also determines the translation of the distribution. The covariance matrix $C$, positive definite and symmetric, determines the shape of the distribution and its graphical interpretation: it defines an iso-density ellipsoid $\{ x_i \in \mathcal{R} | (x - m)^T C^{-1} (x - m) = 1\}$. Through the classic evolutionary setup and by making clever use of population statistics, the algorithm proceeds from generation to generation of sampling until some convergence criteria has been reached.

\subsection{T-HOO}
%$\bm{x^{C_l}_{\mathit{min}}} = \bm{x^{C}_{\mathit{min}}} $,
Recently, a family of algorithms based on discretizing the search domain has been proposed. Only one of these algorithms is going to be discussed here, Truncated Hierarchical Optimistic Optimisation (T-HOO)~\cite{hoo2011}. The T-HOO algorithm was built from the ground up to perform optimisation of noisy functions. It is an iterative, anytime algorithm, which is run for $n_0$ iterations.  The algorithm is based on a number of components. The first one is a data structure called a node and designated $\nu$. The node includes the maximum and minimum values of our search space, vectors $\bm{x_{\mathit{max}}}$ and $\bm{x_{\mathit{min}}}$ respectively. The node also has a property called $\mathit{count}$, which initially is set to zero and a property called $\mathit{sum}$, which, again, is initially set to zero. An unvisited node (i.e. $\mathit{count}=0$) is called a leaf node.  ``Splitting'' a node is the procedure of creating two children $C(\nu) = \{C_l(\nu),C_r(\nu)\}$. Both children are initialised $\bm{x_{\mathit{max}}}$ and $\bm{x_{\mathit{min}}}$ to their parents' nodes values. When there are multiple dimensions to choose from, at  at $k$ according to some random distribution. Thus, for child $C_l(\nu)$ ,   $\bm{x^{C_l}_{\mathit{max,k}}} =\bm{x^{C}_{\mathit{min,k}}} +  \left(\bm{x^{C}_{\mathit{min,k}}} - \bm{x^{C}_{\mathit{max,k}}}\right)/2 $. For the other child,  $\bm{x^{C_r}_{\mathit{min,k}}} = \bm{x^{C}_{\mathit{min,k}}} +  \left(\bm{x^{C}_{\mathit{min,k}}} - \bm{x^{C}_{\mathit{max,k}}}\right)/2 $. In other words, the father node is split into two children nodes, choosing which dimension to split uniform randomly (or according to some other criterion). Each node also has an associated value called $B(\nu)$, which is defined recursively as:  $B(\nu) = min \{U(\nu),max \{B(C_l(\nu)),B(C_r(\nu))\}\}$. Part of this equation is the term $U(\nu)$, which is defined as in Equation \ref{eq:U}:

@@

The MDP defines a single agent environment, fully observable to the agent. In an MDP, the Markov property holds (hence the name), which means that all the information an agent needs in order to act is embedded in the current state.
A possible route of action an agent might take is known as the \textit{policy} $\pi$. A policy is a probabilistic mapping between state and actions, $\pi:S \times A \rightarrow [0,1]$, thus $\pi(s,a) \in [0,1]$,  $\sum\limits_{a\in A}{\pi(s,a)} = 1$. The set of all policies is denoted as $\Pi$. The goal of an agent in an MDP environment is to maximise its long term value. Note that since here we are dealing with continuous actions and states, sets $A$ and $S$ are metric (and infinitive). Also note that we are NOT dealing with continuous time MDPs in this paper, although there is a notion of time. An example of the effects of the discount factor on rewards can be found in Figure \ref{fig:gamma}.

@@

Another point of interest is planning, and when it happens. In open loop planning (and what is usually termed ``planning'' without any further qualifications), an agent forms a plan once and follows it until the end.

@@

A generative model is nothing more than a simulator of future events. This is a less of a requirement than knowing the full model (which entails knowing the full function $T$), but still requires the agent to have some knowledge of the world. This category of algorithms, which are broadly termed ``Rolling Horizon'', has the following steps. From the current state, a number of simulations of possible futures is run by executing a set of action sequences (all of this in the ``mind'' of the agent). The simulations last for a certain number of steps (hence the ``rolling horizon'') or continue until an end state is encountered (i.e. when the MDP has ``absorbing'' states that cannot be acted from).

@@

In standard UCT, a procedure similar to temporal difference learning is followed (or, more closely, monte carlo methods), where one learns the value of each state. An agent then forms a policy by choosing a value greedily. In our case, because our state space has continuous states and actions, this is impossible without the use of a function approximator (or an infinite discretization of space), which is not attempted here. What is attempted though is to learn the policy function directly without reference to state, while planning continuously from the current state. So while the agent forms open-loop plans at each clock tick, in reality it is executing a closed-loop policy. While the strategy is not provably optimal, it can work well in practice~\cite{weinstein2012bandit}. This will be seen by our experimental results (see Section \ref{sec:Experiments}).

@@

\section{Methodology} \label{sec:methodology}
Our methodology is simple, yet efficient. We test T-HOLOP and EVO-P in two simple benchmark problems (see below for more) and Lunar Lander, a classic arcade game. EVO-P is simply CMA-ES used in a rolling horizon setting, while T-HOLOP is T-HOO used in the same manner.

@@

For direct comparison with recent work on sample based planning using trees, in particular
Weinstein and Littman~\cite{weinstein2012bandit} and Pazis and Lagoudakis \cite{PazisBAS},
we used the same problems: double integrator, and inverted pendulum.  The basic
experimental setup followed Weinstein and Littman~\cite{weinstein2012bandit}
and Pazis and Lagoudakis \cite{PazisBAS}
closely, except that they used a fixed noise level for each experiment whereas we varied the
noise level to explore how performance was affected.

@@

The problem becomes harder as the time interval $\delta_t$ at %word missing?
each action is applied increases,
and following Pazis and Lagoudakis we set $(\delta_t = 0.5s)$.

@@

\subsection{Lunar Lander}
In this section we give a brief overview of the Game of Lunar Lander, a popular arcade game.

@@

The spaceship within this game world is modelled as an a circular mass with some basic physical properties, such as a position within the game world $\bm {s}$, a velocity $\bm {v}$, an orientation $\bm{d}$, an angular velocity $\omega$ and a radius of a bounding sphere $r$ used for collision detection with the landscape.

@@

The ship colliding with the landscape constitutes the end of the {\itshape Lunar Lander} game. The conditions surrounding this collision, including speed, orientation of the ship constitute whether the nature of the collision is a success or a failure. In this experimental setup with accept a velocity of $5$ and an angle of $< 0.1$ radians as successful landing criteria.

An agent can modify the ship directions by applying linear thrust $\bm{a_T}$ and the angular impulse $\bm{a_I}$. The controllers have to identify a vector of real numbers $\bm{A}$, which each element of the set $a_i \in [0,1]$. Each even element of the vector represents the linear thrust, whereas each odd element represents angular velocity. Each linear thrust vector element is transformed by applying as follows:

@@

Lunar Lander has a setup where an agent gets rewards if and only if it manages to actually land the ship on the pad. This creates the problem of rollouts (whether used by EVO-P or T-HOLOP) rarely will see a reward. In order to circumvent this, as it is common in other games, we use a heuristic for each state. The heuristic was partially inspired by the two-stage approach found here : \url{http://www.cs.berkeley.edu/~jrs/4/lunar.html}\cite{lunar}. Initially, we need define a number of vectors that portray ship directions:

@@

The vectors defined in Equation \ref{eq:vectors} define three directions for the ship. The first one, $\bm{du}$, define a ship standing upright, the second one, $\bm{dl}$, define tilting to the left, while the third one $\bm{dr}$ a ship tilting to the right. We make use of these vectors in order to help align the ship in the right direction, depending on which way the landing point is.

@@

In the above Equation \ref{eq:h1} the constant C denotes the maximum magnitude of speed, which we set to $C=30$. $D = 60$ is the distance from the landing point $np$ where this heuristic takes precedence.  Intuitively, the upper part of the heuristic $h$ tries to get the ship above the landing point as fast as possible while maintaining a certain speed. The second part of the heuristic tries to minimise velocity on the $x$ axis (as the ship tries to land with maximum speed $C_l = 2$), maintain a minimum landing speed on the $y$ axis and get a certain angle (the summation term $f$) as close to vector $du$, i.e. help the ship stay upright through the duration of the rollout.  It's worth noting the main reason why we use a heuristic of this kind, one that includes not only the final state, but makes use of the path of the state as well, is that there is not enough information to discern states if one takes into account ship direction only; without incorporating the angular velocity,  the ship might be spinning madly, but the end of the rollout might find it in the correct direction. Finally notice that the heuristic $h$ gets the value $p=10000$ if the ship has crashed at any point during the rollout.

@@

 EVO-P got as score of $-41.69$. It also  seems to have a drop in performance as we increase function evaluations, presumably because it converges to a false minima without any uncertainty handling mechanism to stop this. This kind of behaviour has been observed previously in Evolutionary Algorithms~\cite{hansen2009tec} when having noisy function.

For the double integrator experiment, we had slightly different results. The Weinstein paper reports a mean of $-2.72$ for 200 iterations. Using T-HOLOP we got a better mean of $-2.62$. A far better result of $-1.74$ is obtained by EVO-P.

@@

\begin{table}[h]
	\begin{tabular}{lrrrr}
\hline
 Map Name   &   Success Mean &   Success CI &   Ticks Mean &   Tick CI \\
\hline
 Map A       &            1   &     0        &       2925.9 &   $\pm 184.437$ \\
 Map B  &            1   &     0        &       2733   &   $\pm 159.34$  \\
 Map C  &            1   &     0        &       2293   &   $\pm 123.917$ \\
 Map D  &            0.9 &     $\pm 0.226216$ &       2421.6 &   $\pm 682.893$ \\
\hline
\end{tabular}
\label{tb:results}
\caption{EVO-P results for different lunar lander maps}
\end{table}

For lunar lander, we do $10$ runs for each map in Figure \ref{fig:maps}, and the results for EVO-P can be seen in Table I. The results include the mean number of successful landings (\emph{Success Mean}) and the mean time it took to land the ship (\emph{Ticks Mean}), alongside their confidence for the $95$th interval.   For EVO-P we used the default CMA-ES population size, each individual was bound between, each gene had a length of $8$ and each action pair ($x[i], x[i+1]$) was repeated $7$ times (where $x[i]$ represents thrust and $x[i+1]$ angular thrust). We derived these numbers experimentally and we did not optimise for them (not even implicitly) to any serious degree.

There are a number of things to notice here. The first one is that the only in Map D, where the ship starts the problem facing a wall do we get a failure to land, and this is due to a crash. This has to do with the fact that even a minor thrust can push you a bit forward and drive the ship to a position where there is no longer the possibility of recovery from a crush (or the combination of that recovery is extremely hard to find). The second thing to observe is that we get there slowly, but this is by design, as we he have made the flying speed of the ship controllable. One can potentially tweak all the default parameters we have used, plus the ones incorporated in CMA-ES as a kind of meta-heuristic search and achieve much better results. Another thing to note is that EVO-P lands the ship (on average) almost at the same time on MAPS B, C, D, although map has the same landscape as the other two but the ship is much closer. This is due to the fact that the minimum landing speed, set on one of our heuristics, is really low. The final thing we would like to note is that we have only included the results for EVO-P because T-HOLOP failed to land the ship using the same heuristic, and behaved badly. It's worth noting here that T-HOLOP requires rewards between certain bounds, so they can be normalised around $[0,1]$ and this is not the case with our heuristic. For completeness purposes, we did change the UCB-like node selection policy of T-HOO to e-greedy, in which case we \emph{did} get the ship to behave reasonably, albeit the failure to land was evident here as well. This is important we think for a number of reasons. The first one is that it shows that different Heuristics might be needed for different algorithms. Although this might be self-evident, it shows that one must design heuristics with a specific algorithm in mind. The second is, but we can speculate at the time, that T-HOLOP spends too much trying to solve a hard problem than the one at hand. It will converge eventually, but the the Guassian approximation of CMA-ES seems to outperform it in the short term.

>>>>

\begin{abstract}
It is often the case that games have continuous dynamics and allow for continuous actions, possibly with with some added noise. For larger games with complicated dynamics, having agents learn offline behaviours in such a setting is a daunting task. On the other hand, provided a forward model is available, one might try to spread the cost of search/learning in a rolling horizon fashion (e.g. as in Monte Carlo Tree Search). In this paper we compare T-HOLOP (Truncated Hierarchical Open Loop Planning), an open loop planning algorithm at least partially inspired by MCTS, with a version of evolutionary planning that uses CMA-ES (which we call EVO-P) in two planning benchmark problems (Cart Balancing and the Double Integrator) and in Lunar Lander, a classic arcade game. We show that EVO-P outperforms T-HOLOP in the classic benchmarks, while T-HOLOP is unable to find a solution using the same heuristics. We conclude that off-the-shelf evolutionary algorithms can be used successfully in a rolling horizon setting, and that a different type of heuristics might be needed under different optimisation algorithms.

@@

\section{Introduction} \label{Introduction}
One of the most common uses of function optimisation is solving control problems. If an agent has access to a generative model of the world (i.e. a simulator) in which it is going to act, a control problem becomes a search/optimisation problem, usually referred to as ``simulation based planning''. Recently, a family of algorithms, mostly known as Monte Carlo Tree Search (MCTS)~\cite{browne2012survey}, has been applied to planning problems of discrete states and actions with considerable success. The focus of these algorithms is to attack extremely large state spaces of perfect information games, where one can sample rewards from the state space easily (e.g. Computer Go~\cite{gelly2006modification}). With access to a generative model (from which one can easily sample) and perfect sensor information, the following procedure works well: the agent receives sensor information, formulates a plan of action, performs the first action of that plan, receives a new state, re-formulates a plan, acts again (using the first action of its plan),  ad infinitum.

@@

Recently, algorithms stemming from, or at least partially inspired by, MCTS aiming at solving planning problems involving continuous states and actions have come to light. A typical such algorithm is Hierarchical Optimistic Open Loop Planing (HOLOP)~\cite{weinstein2012bandit}. HOLOP is based partially on a strong~\cite{hoo2011} real-valued optimisation algorithm called HOO (for ``Hierarchical Optimistic Optimisation''). In this paper two contributions are made. The first contribution is the comparison of two versions of a very popular and strong evolutionary strategy: EVO-P, the open loop planning algorithm using Covariance Matrix Adaptation Evolutionary Strategy (CMA-ES)~\cite{hansen2003reducing} and a version of HOLOP, which stands for Hierarchical Open Loop Optimistic Planning~\cite{hoo2011} on two standard benchmark problems, the double integrator and the reverse pendulum.  For speed purposes, we use T-HOO (Truncated Hierarchical Optimistic Optimisation), a faster version of HOO~\cite{hoo2011}, as the engine of HOLOP, hence we name the algorithm T-HOLOP.  Our second contribution is the use of these algorithms in Lunar Lander, a popular arcade game, coupled with a strong two-stage heuristic. We show that while EVO-P performs reasonably well, T-HOLOP struggles.

While evolution is commonly used to evolve a reactive neural network controller (in a process commonly called neuro-evolution) for these and other RL problems,
it should be noted that evolution is being applied in a very different way here. In this paper, evolution is applied to perform each action given the current state.  This approach can only be applied when a generative
model is available, but has the advantage of offering immediate good performance without any
prior learning.  The disadvantage, compared to neuro-evolution, is that every action performed requires
CPU time for the simulations.  This is described in more detail in Section \ref{sec:Planning} and in line with what was done in Perez et al. \cite{perez2013rolling}, although this time in a continuous setting.

@@

This is the first background section of this paper and aims to introduce the two core optimisation algorithms on which our contributions are based. We will present both algorithms from an implementer's viewpoint, without explicitly justifying the design and mathematical choices behind them, nor going into too much detail. Interested readers should look at the cited bibliography for extensive discussions.

In general, the optimisation problem can be defined as: Given a function $f:\bm{X}\rightarrow Y $, where $X \in \mathcal{R}^n$, with $n$ being the length of vector $\bm{X}$. The domain $\bm{X}$ is usually called the search space. Our goal is to find a vector $\bm{x}_0$, for which $f(\bm{x}) > f(\bm{x_0}) $ for all $x$ (also known as ``minimisation'') or $f(\bm{x}) < f(\bm{x_0})$ for all $x$ (known as ``maximisation'').
There are an enormous number of algorithms devoted to this problem, but only two will be presented and compared here. Both of these are considered state of the art in their respective fields and do not require the calculation of a gradient. If one has an analytic expression of the gradient or if the functions proposed are smooth enough, it might be better to look elsewhere for solutions (e.g. Conjugate Gradient Methods~\cite{shewchuk1994introduction}).

@@

CMA-ES creates a population of individuals by sampling from an MND: $\mathcal{N}(m,C)$, which is uniquely defined by the distribution mean $m \in \mathcal{R}^{n}$ and its covariance matrix $C \in \mathcal{R}^{n \times n}$. The top of the density function, which corresponds to $m$, also determines the translation of the distribution. The covariance matrix $C$, positive definite and symmetric, determines the shape of the distribution and its graphical interpretation: it defines an iso-density ellipsoid $\{ x_i \in \mathcal{R} | (x - m)^T C^{-1} (x - m) = 1\}$. Through the classic evolutionary setup, and by making clever use of population statistics, the algorithm proceeds from generation to generation of sampling until some convergence criteria has been reached.

\subsection{T-HOO}
%$\bm{x^{C_l}_{\mathit{min}}} = \bm{x^{C}_{\mathit{min}}} $,
Recently, a family of algorithms based on discretizing the search domain has been proposed. Only one of these algorithms is going to be discussed here, Truncated Hierarchical Optimistic Optimisation (T-HOO)~\cite{hoo2011}. The T-HOO algorithm was built from the ground up to perform optimisation of noisy functions. It is an iterative, anytime algorithm, which is run for $n_0$ iterations.  The algorithm is based on a number of components. The first one is a data structure called a node and designated $\nu$. The node includes the maximum and minimum values of our search space, vectors $\bm{x_{\mathit{max}}}$ and $\bm{x_{\mathit{min}}}$ respectively. The node also has a property called $\mathit{count}$, which initially is set to zero and a property called $\mathit{sum}$, which, again, is initially set to zero. An unvisited node (i.e. $\mathit{count}=0$) is called a leaf node.  ``Splitting'' a node is the procedure of creating two children $C(\nu) = \{C_l(\nu),C_r(\nu)\}$. Both children are initialised $\bm{x_{\mathit{max}}}$ and $\bm{x_{\mathit{min}}}$ to their parents' node values when there are multiple dimensions to choose from, at $k$ according to some random distribution. Thus, for child $C_l(\nu)$ ,   $\bm{x^{C_l}_{\mathit{max,k}}} =\bm{x^{C}_{\mathit{min,k}}} +  \left(\bm{x^{C}_{\mathit{min,k}}} - \bm{x^{C}_{\mathit{max,k}}}\right)/2 $. For the other child,  $\bm{x^{C_r}_{\mathit{min,k}}} = \bm{x^{C}_{\mathit{min,k}}} +  \left(\bm{x^{C}_{\mathit{min,k}}} - \bm{x^{C}_{\mathit{max,k}}}\right)/2 $. In other words, the father node is split into two children nodes, choosing which dimension to split uniform random (or according to some other criterion). Each node also has an associated value called $B(\nu)$, which is defined recursively as:  $B(\nu) = min \{U(\nu),max \{B(C_l(\nu)),B(C_r(\nu))\}\}$. Part of this equation is the term $U(\nu)$, which is defined as in Equation \ref{eq:U}:

@@

The MDP defines a single agent environment, fully observable to the agent. In an MDP, the Markov property holds (hence the name), which means that all the information an agent needs in order to act is embedded in the current state.
A possible route of action an agent might take is known as the \textit{policy} $\pi$. A policy is a probabilistic mapping between state and actions, $\pi:S \times A \rightarrow [0,1]$, thus $\pi(s,a) \in [0,1]$,  $\sum\limits_{a\in A}{\pi(s,a)} = 1$. The set of all policies is denoted as $\Pi$. The goal of an agent in an MDP environment is to maximise its long term value. Note that, since we are dealing with continuous actions and states here, sets $A$ and $S$ are metric (and infinitive). Also note that we are NOT dealing with continuous time MDPs in this paper, although there is a notion of time. An example of the effects of the discount factor on rewards can be found in Figure \ref{fig:gamma}.

@@

Another point of interest is \textit{when} planning happens. In open loop planning (and what is usually termed ``planning'' without any further qualifications), an agent forms a plan once and follows it until the end.

@@

A generative model is nothing more than a simulator of future events. This is less of a requirement than knowing the full model (which entails knowing the full function $T$), but still requires the agent to have some knowledge of the world. This category of algorithms, which are broadly termed ``Rolling Horizon'', has the following steps. From the current state, a number of simulations of possible futures is run by executing a set of action sequences (all of this in the ``mind'' of the agent). The simulations last for a certain number of steps (hence the ``rolling horizon'') or continue until an end state is encountered (i.e. when the MDP has ``absorbing'' states that cannot be acted from).

@@

In standard UCT, a procedure similar to temporal difference learning is followed (or, more closely, Monte Carlo methods), where one learns the value of each state. An agent then forms a policy by choosing a value greedily. In our case, because our state space has continuous states and actions, this is impossible without the use of a function approximator (or an infinite discretization of space), which is not attempted here. What is attempted though is to learn the policy function directly without reference to state, while planning continuously from the current state. So while the agent forms open-loop plans at each clock tick, in reality it is executing a closed-loop policy. While the strategy is not provably optimal, it can work well in practice~\cite{weinstein2012bandit}. This will be seen by our experimental results (see Section \ref{sec:Experiments}).

@@

\section{Methodology} \label{sec:methodology}
Our methodology is simple, yet efficient. We test T-HOLOP and EVO-P in two simple benchmark problems (see below for more detail) and n Lunar Lander, a classic arcade game. EVO-P is simply CMA-ES used in a rolling horizon setting, while T-HOLOP is T-HOO used in the same manner.

@@

For direct comparison with recent work on sample based planning using trees, in particular
Weinstein and Littman~\cite{weinstein2012bandit} and Pazis and Lagoudakis \cite{PazisBAS},
we used the same problems: double integrator and inverted pendulum.  The basic
experimental setup followed Weinstein and Littman~\cite{weinstein2012bandit}
and Pazis and Lagoudakis \cite{PazisBAS}
closely, except that they used a fixed noise level for each experiment whereas we varied the
noise level to explore how performance was affected.

@@

The problem becomes harder as the time interval $\delta_t$ that is applied at each action, increases,
and following Pazis and Lagoudakis we set $(\delta_t = 0.5s)$.

@@

\subsection{Lunar Lander}
In this section we give a brief overview of the game of Lunar Lander, a popular arcade game.

@@

The spaceship within this game world is modelled as a circular mass with some basic physical properties, such as a position within the game world $\bm {s}$, a velocity $\bm {v}$, an orientation $\bm{d}$, an angular velocity $\omega$ and a radius of a bounding sphere $r$ used for collision detection with the landscape.

@@

The ship colliding with the landscape constitutes the end of the {\itshape Lunar Lander} game. The conditions surrounding this collision, including speedand orientation of the ship, constitute whether the nature of the collision is a success or a failure. In this experimental setup we accept a velocity of $5$ and an angle of $< 0.1$ radians as successful landing criteria.

An agent can modify the ship's direction by applying linear thrust $\bm{a_T}$ and angular impulse $\bm{a_I}$. The controllers have to identify a vector of real numbers $\bm{A}$, with each element of the set $a_i \in [0,1]$. Each even element of the vector represents the linear thrust, whereas each odd element represents angular velocity. Each linear thrust vector element is transformed by applying as follows:

@@

Lunar Lander has a setup where an agent gets rewards if and only if it manages to actually land the ship on the pad. This creates the problem that rollouts (whether used by EVO-P or T-HOLOP) will rarely see a reward. In order to circumvent this, as it is a problem common to other games as well, we use a heuristic for each state. The heuristic was partially inspired by the two-stage approach found here : \url{http://www.cs.berkeley.edu/~jrs/4/lunar.html}\cite{lunar}. Initially, we need to define a number of vectors that portray ship directions:

@@

The vectors defined in Equation \ref{eq:vectors} define three directions for the ship. The first one, $\bm{du}$, defines a ship standing upright, the second one, $\bm{dl}$, defines it tilting to the left, while the third one $\bm{dr}$ tilting to the right. We make use of these vectors in order to help align the ship in the right direction, depending on where the landing point is.

@@

In the above Equation \ref{eq:h1} the constant C denotes the maximum magnitude of speed, which we set to $C=30$. $D = 60$ is the distance from the landing point $np$, where this heuristic takes precedence.  Intuitively, the upper part of the heuristic $h$ tries to get the ship above the landing point as fast as possible while maintaining a certain speed. The second part of the heuristic tries to minimise velocity on the $x$ axis (as the ship tries to land with maximum speed $C_l = 2$), maintain a minimum landing speed on the $y$ axis and get a certain angle (the summation term $f$) as close to vector $du$ as possible, i.e. help the ship stay upright through the duration of the rollout.  It's worth noting that the main reason why we use a heuristic of this kind (one that includes not only the final state, but makes use of the path of the state as well) is that there is not enough information to discern states if one only takes into account ship direction.  Without incorporating the angular velocity,  the ship might be spinning madly, but the end of the rollout might find it in the correct direction. Finally, notice that the heuristic $h$ gets the value $p=10000$ if the ship has crashed at any point during the rollout.

@@

 EVO-P got a score of $-41.69$. It also seems to experience a drop in performance as we increase function evaluations, presumably because it converges to a false minima without any uncertainty handling mechanism to stop this. This kind of behaviour has been observed previously in Evolutionary Algorithms~\cite{hansen2009tec} when having noisy function.

For the double integrator experiment, we had slightly different results. The Weinstein paper reports a mean of $-2.72$ for 200 iterations. Using T-HOLOP we got a better mean of $-2.62$. A far better result of $-1.74$ was obtained by EVO-P.

@@

\begin{table}[h]
	\begin{tabular}{lrrrr}
\hline
 Map Name   &   Success Mean &   Success CI &   Ticks Mean &   Tick CI \\
\hline
 Map A       &            1   &     0        &       2925.9 &   $\pm 184.437$ \\
 Map B  &            1   &     0        &       2733   &   $\pm 159.34$  \\
 Map C  &            1   &     0        &       2293   &   $\pm 123.917$ \\
 Map D  &            0.9 &     $\pm 0.226216$ &       2421.6 &   $\pm 682.893$ \\
\hline
\end{tabular}
\label{tb:results}
\caption{EVO-P results for different Lunar Lander maps}
\end{table}

For Lunar Lander, we perform $10$ runs for each map in Figure \ref{fig:maps}, and the results for EVO-P can be seen in Table I. The results include the mean number of successful landings (\emph{Success Mean}) and the mean time it took to land the ship (\emph{Ticks Mean}), alongside the confidence for the $95$th interval.   For EVO-P we used the default CMA-ES population size, each individual gene bound between $[0,1]$, each gene had a length of $8$ and each action pair ($x[i], x[i+1]$) was repeated $7$ times (where $x[i]$ represents thrust and $x[i+1]$ angular thrust). We derived these numbers experimentally and did not optimise for them (not even implicitly) to any serious degree.

There are a number of things to notice here. The first is that only in Map D, where the ship starts the problem facing a wall, do we get a failure to land, and this is due to a crash. This has to do with the fact that even a minor thrust can push you a bit forward and drive the ship to a position where there is no longer the possibility of recovery from a crush (or the combination of that recovery is extremely hard to find). The second thing to observe is that we get there slowly, but this is by design, as we he have made the flying speed of the ship controllable. One can potentially tweak all the default parameters we have used, plus the ones incorporated in CMA-ES, as a kind of meta-heuristic search and achieve much better results. Another thing to note is that EVO-P lands the ship (on average) almost at the same time on MAPS B, C, D, although map B has the same landscape as the other two but the ship is much closer. This is due to the fact that the minimum landing speed, set on one of our heuristics, is really low. The final thing we would like to point out is that we have only included the results for EVO-P because T-HOLOP failed to land the ship using the same heuristic, and behaved badly. It's worth noting here that T-HOLOP requires rewards between certain bounds, so they can be normalised around $[0,1]$ and this is not the case with our heuristic. For completeness purposes, we did change the UCB-like node selection policy of T-HOO to e-greedy, in which case we \emph{did} get the ship to behave reasonably, albeit the failure to land was evident here as well. This, we think, is important for a number of reasons. The first one is that it shows that different Heuristics might be needed for different algorithms. Although this might be self-evident, it shows that one must design heuristics with a specific algorithm in mind. The second is, but we can speculate only at this time, that T-HOLOP spends too much effort trying to solve a hard problem than the one at hand. It will converge eventually, but the the Gaussian approximation of CMA-ES seems to outperform it in the short term.
